Satisfiability Modulo Theories Competition (SMT-COMP) 2010: Rules and Procedures

### 2015-02-11 rules/rules10.tex
svn propdel svn:executable

\title{Satisfiability Modulo Theories Competition (SMT-COMP) 2011: Rules and 
Procedures
}

@@

\author{
Roberto Bruttomesso\\
Computer Science\\
USI, Lugano (Switzerland)
\and
Morgan Deters \\
Computer Science \\
New York University \\
\and
Alberto Griggio\\
ES Division\\
FBK, Trento (Italy)
}

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{PSS02,leberre+03}.  More information
on the history and motivation for SMT-COMP can be found at the
SMT-COMP web site, \url{www.smtcomp.org}, and in reports on previous
competitions~(\cite{SMTCOMP-2008,BDOS08,SMTCOMP-FMSD,SMTCOMP-JAR}).
SMT-COMP 2011 is affiliated with the 23nd International
Conference on Computer Aided Verification~(CAV).

The rest of this document, updated from the last year's
version\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, and Aaron Stump.},
describes the rules and competition procedures for SMT-COMP~2011.
The principal change from last year's version is:

\begin{itemize}
\item the introduction of an ``application track,'' based on incremental
      model-checking queries, to promote the integration of SMT solvers
      within actual verification frameworks;
%\item updates to the benchmark selection algorithm, to reduce
%      overwhelming a competition division with a single ``family'' of
%      benchmarks.
% rb 25/3: see later
%\item updates to the difficulty calculation;
%\item a track for parallel solvers, to run before the regular competition.
%\item 5 random ``check'' benchmarks will be generated using Robert
%  Brummayer's SMT fuzzing tool, after all the solvers are uploaded.
%  These will be included in the competition, to help promote attention
%  to the problem of solver correctness (cf.~\cite{brummayer+09}).
\end{itemize}

@@

\header{Solver format.} %
An entrant to SMT-COMP is an SMT solver
submitted using the SMT-Exec service.  SMT-Exec enables members of the
SMT research community to run solvers on jobs consisting of benchmarks
from the SMT-LIB benchmark library.  Jobs are run on a 12-node computer
cluster purchased with funds from a National Science Foundation
Computing Research Infrastructure grant.  SMT-Exec is provided free of
charge, but it does require a minimal registration, which verifies an
email address and prevents misuse of the service.  Registered users
may then upload their own solvers to run, or may run public solvers
already uploaded to SMT-Exec.  SMT-Exec provides a variety of tabular
and graphical displays of results of solver executions, for
comparison. For the main (sequential-solver)
% rb 23/3: I guess also application track runs on 
% the same clusters as sequential
and application tracks 
of SMT-COMP 2011, SMT-Exec will be configured so that jobs are run on a 64-bit,
uniprocessor Linux kernel.  For the parallel-solver track, SMT-Exec
will be configured so that jobs are run on a 64-bit, multiprocessor
Linux kernel.\footnote{This year, the competition may take
  advantage of off-site computing resources not normally part of
  SMT-Exec; please stay tuned to the SMT-COMP mailing list for final
  details regarding the effective SMT-COMP 2011 competition cluster.}

@@

\subsection*{Deadlines} %

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via SMT-Exec by 7pm, Eastern U.S.
time, July~9th, 2011.  At that time the SMT-Exec service will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until 7pm, Eastern U.S. time, July~11th, 2011.  We strongly encourage
participants to use this two day grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using SMT-Exec after the original deadline on July~9th.

@@

\paragraph{Application track.} %
The same deadlines and procedure for submitting to the main track will be used
also for the application track. Submissions to both the application track and
the main competition are independent: participants must submit explicitly to
both events to participate in both, and they may submit different (or
differently configured) solvers to each.  Benchmarks will be scrambled also
for this division, using the same scrambler and seed as the main track.
Entrants should still include a system description, as for the regular
competition.

\paragraph{Parallel solver track.} %
Solvers employing concurrency are invited to participate in a
demonstration that will be run before the regular competition.
Parallel solvers must be submitted by the same July~9th deadline,
with the same two-day grace period for resubmission as in the regular
competition.  Submissions to both the parallel solver track
and the main competition are independent: participants must submit
explicitly to both events to participate in both, and they may submit
different (or differently configured) solvers to each.  Benchmarks may
or may not be scrambled for this division.  Entrants should still
include a system description, as for the regular competition.

\medskip
\noindent
Note that the SMT-COMP organizers are considering ways to include
other (``historical'' or otherwise relevant) solvers in all the competition tracks for
demonstration and comparison.  The organizers reserve the right to
include or exclude such solvers, and to make simple modifications
to historical sequential solvers to (na\"ively) take advantage of
multiple processors 
or to connect them with a parser for the SMT-LIB version 2.0 format
(\eg\ by implementing a wrapper tool).

@@

\subsection{Main track and parallel track}
\label{sec:exec:main}

\header{Dates of competition.}
%
We anticipate that the bulk of the main track of the competition will take place during
the course of CAV 2011, from July~14th to~20th.  Results will be
announced in a special session of CAV, on the last day of the
conference, as well as on the SMT-COMP web site.  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.
%
If there are enough competitors to run the parallel track, we intend
to run it before or after the main competition (depending on the
estimated duration of the track).

\header{Input and Output.}
%
Participating solvers must read a single benchmark script (defined
below, not part of the 2.0 standard), presented on its standard input
channel. The script is in the concrete syntax of the SMT-LIB format,
version 2.0.  A benchmark script is essentially just the translation
of a benchmark from the version 1.2 specification.  In more detail, a
\textbf{benchmark script} is just a script where:

\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command.
% rb 25/3: added
% \item A (single) \akey{set-option :print-success false} command, sent just
%       after $\akey{set-logic}$. This is to
%       avoid the output ``\texttt{success}'' to interfere with the
%       answer of the solver to the benchmark. 
\item The $\akey{exit}$ command is the last command.
\item There is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item There is at most one $\akey{set-info}$ command for \texttt{status}.
\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item Extra symbols are declared exactly once before any
  use, using $\akey{declare-sort}$
  %, $\akey{define-sort}$,
  %$\akey{declare-fun}$, or $\akey{define-fun}$.  
  or $\akey{declare-fun}$.
  They must be part of the allowed signature expansion for the logic.
  Moreover, all sorts declared with a $\akey{declare-sort}$ command must have zero arity.
\item No other commands besides the ones just mentioned may be used.
\end{enumerate}

% \noindent Note that the final version of the 2.0 format will be frozen
% in early March.  
\noindent The SMT-LIB format specification is publicly
available from the ``Documents'' section of the SMT-LIB
website~\cite{SMT-LIB}.  Solvers will be given formulas just from the
Problem Divisions indicated during their submission to SMT-Exec.
Example benchmark scripts for several Problem Divisions are reported in the Appendix.
Note that they are provided for illustrative purposes only: 
please refer to the SMT-LIB format specification 
and the above definition of benchmark script for the official specification 
of the input format for SMT-COMP.

\medskip
Each SMT-COMP entrant is then expected to attempt to report on its
standard output channel whether the formula is satisfiable
(``\texttt{sat}'', in lowercase, without the quotation marks) or unsatisfiable
(``\texttt{unsat}'').  An entrant may also report ``\texttt{unknown}''
to indicate that it cannot determine satisfiability of the formula.
For more detailed information on the output format, see the
description on the SMT-Exec ``Upload a Solver'' page.

@@

\header{Aborts and unparsable output.} %
Solvers which exit before the time
limit without reporting a result (\ie\ due to exhausting memory, crashing,
or producing output other than \texttt{sat}, \texttt{unsat}, or
\texttt{unknown})
will be considered to have aborted. 
An abort scores the same as if the output is ``\texttt{unknown}''. 
% rb 25/3: added
Also, as a further measure to prevent misjudgments of solvers,
the first burst of  ``\texttt{success}'' outputs will be 
ignored\footnote{
Note that SMT-LIBv2 requires to produce a ``\texttt{success}'' answer
after each \akey{set-logic}, \akey{declare-sort}, \akey{declare-fun} and
\akey{assert} command (among others), unless the option
\akey{:print-success} is set to false: ignoring the first burst of
\texttt{success} outputs therefore allows for submitting fully-compliant
solvers without the need of a wrapper script, while still allowing entrants
of SMT-COMP 2010 to run without changes.}.

@@

\subsection{Application track}
\label{sec:exec:application}

The application track aims at evaluating SMT solvers when interacting
with an external verification framework, \eg, a model
checker. This interaction, ideally, happens by means of an online
communication between the model checker and the solver: the model
checker repeatedly sends queries to the SMT solver, which in turn
answers either sat or unsat.  In this interaction an SMT-solver is
required to accept queries incrementally via its standard input channel.

In order to facilitate the evaluation of the solvers in this track, we
propose to set up a ``simulation'' of the aforementioned interaction. In
particular each benchmark in the application track represents a realistic
communication trace, containing multiple \akey{check-sat} commands (possibly
with corresponding \akey{push 1}/\akey{pop 1} commands), which
is parsed by a {\em trace executor}. The trace executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to the SMT solver
      (through stdin);
\item it prevents ``look-ahead'' behaviours of SMT solvers;
\item it records time and answers for each call, possibly aborting the execution
      in case of a wrong answer;
\item it guarantees a fair execution for all solvers by abstracting from any possible
      crash, misbehaviour, etc. that may happen on the model checker side.
\end{itemize}

\header{Input and Output.}
Participating solvers will be connected to a trace executor 
which will incrementally send commands to the standard input channel of the solver
and read responses from the standard output channel of the solver.
The commands will be taken from an \textbf{incremental benchmark script},
which is an SMT-LIB 2.0 script which satisfies points 1., 2., 4.--7. of the definition of benchmark script given in \S\ref{sec:exec:main}, plus the following:

\begin{enumerate}
\item[3'.] There are one or more \akey{check-sat} commands, 
  each preceded by one or more \akey{assert} commands 
  and zero or more \akey{push 1} commands, 
  and followed by zero or more \akey{pop 1} commands.
\end{enumerate}

\noindent
Furthermore, the trace executor will send a single 
\akey{set-option :print-success true} command to the solver before 
sending commands from the incremental benchmark script.

\medskip
Solvers must respond immediately to the commands sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification: that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
and with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands.

\header{Timeouts.}
A time limit is set for the whole execution of each application
benchmark (consisting of multiple queries).

\header{Scoring.}
The score for each benchmark is a pair $\langle n,m\rangle$, with
$n\in[0,N]$ an integral number of points scored for the benchmark,
where $N$ is the number of \akey{check-sat} commands
in the benchmark.  $m\in[0,T]$ is the (real-valued) time is seconds, where $T$ is
the timeout.  The score for the benchmark is initialized with
$\langle0,0\rangle$ and then computed as follows.
\begin{itemize}
\item A correctly-reported \texttt{sat} or \texttt{unsat} answer after
  $s$~seconds (counting from the beginning of this particular
  \akey{check-sat}) contributes $\langle1,s\rangle$ to the running
  score.
\item An answer of \texttt{unknown}, an unexpected answer, a crash, or a memory-out during
  execution of the query, or a benchmark timeout, aborts the execution
  of the benchmark and assigns the current value of the running score
  to the benchmark.  (Recall that there is one timeout for the entire
  benchmark; there are no individual timeouts for queries.)
\item The first incorrect answer to a \akey{check-sat} command has the effect of terminating the
  trace executor, and the returned score for the overall benchmark
  will be $\langle0,0\rangle$, effectively canceling the running score.
\end{itemize}

For example, if a benchmark has 5~\akey{check-sat} commands, and a
timeout of 100~seconds, and a solver solves the first four in
10~seconds each, then times out on the fifth, then the solver's score
is $\langle4,40\rangle$.  If another solver solves each of the first
four in 10~seconds each, and the fifth in another 40~seconds, its
score is $\langle5,80\rangle$.  If a third solver solves the first
four queries in 2~seconds each, but incorrectly answers the fifth, its
score is $\langle0,0\rangle$.

As queries are only presented in order, this scoring system may mean
that relatively ``easier'' queries are hidden behind more difficult
ones located at the middle of the query sequence.

Scores are compared lexicographically---a solver with a higher
$n$-value wins, with the cumulative time only used to break ties.
Benchmarks' scores are summed componentwise to form a solver's total
score for the competition.

@@

We expect the problem divisions for SMT-COMP 2011 to include the following
SMT-LIB \emph{logics}.  These logics are specified in SMT-LIB format on the
SMT-LIB web page.  Note that the ``QF\_'' prefix means the division's
formulas are quantifier-free.  However, the organizers reserve the right to add
(remove) divisions if (not) enough benchmarks and solvers exist for a
particular division.

@@

\subsection{Main and parallel tracks}

\header{Benchmark sources.} %
Benchmark formulas for these divisions
will be drawn from the SMT-LIB library.  Any benchmarks added to
SMT-LIB by the May 15th release (see the timeline in
Section~\ref{sec:timeline}) will be considered eligible.  SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.

\header{Benchmark availability.} % A first release of the competition
Benchmarks will be made available within May 15th, 2011.
% A second and almost final release will be available on June 1st. 
No additional
benchmarks will be added after this date, but benchmarks can be
modified or removed to fix possible bugs or other issues, or to adjust their difficulty score (see below). 
The final release that will be used for the competition will be posted on June
30th. The set of selected benchmarks will be published when the
competition begins.

\header{Benchmark demographics.} %
In SMT-LIB, benchmarks are organized according to \emph{families}.  A benchmark
family contains problems that are similar in some significant way.  Typically
they come from the same source or application, or are all output by the same
tool.  \emph{Each top-level subdirectory within a division represents a distinct
family.}  This represents a significant change from
competition years before 2010, where
\emph{terminal} subdirectories delineated families.
%
Each benchmark in SMT-LIB also has a \emph{category}.  There are four possible
categories:
%
\begin{itemize}
\item \emph{check.} These benchmarks are hand-crafted to test whether
  solvers support specific features of each division.  In particular,
  there are checks for integer completeness (\ie\ benchmarks that are
  satisfiable under the reals but not under the integers) and big
  number support (\ie\ benchmarks that are likely to fail if integers
  cannot be represented beyond some maximum value, such as
  $2^{31}-1$).

  \noindent%\textbf{New for 2010:} 
  Using the same random seed as
  for benchmark selection and scrambling, 5 ``check'' benchmarks will
  be randomly generated using Robert Brummayer's SMT fuzzing
  tool~\cite{brummayer+09}.  The (exact version of this) tool will be
  publicly available from the SMT-COMP~2011 web site before the
  competition.  Since these benchmarks are generated after the random
  seed is fixed, they cannot be known in advance to any competitors,
  including the organizers.  The rationale for including these
  benchmarks is to try to take a step towards stronger certification
  that solvers are correct.  In future years, we envision SMT-COMP
  requiring that solvers pass some kind of pre-qualifying round based
  on SMT fuzzing.  The inclusion of these fuzzing benchmarks that are
  not known in advance is intended to encourage (without requring)
  solver implementors to test their solvers using fuzzing or similar
  bug-discovery techniques.
\item \emph{industrial.} These benchmarks come from some real application
      and are produced by tools such as bounded model checkers, static analyzers, extended
      static checkers, etc.
\item \emph{random.} These benchmarks are randomly generated.
\item \emph{crafted.} This category is for all other benchmarks.  Usually,
  benchmarks in this category are designed to be particularly difficult or to
  test a specific feature of the logic.
\end{itemize}

\header{Benchmark selection.} %
Before the selection process, each benchmark will be assigned a
\emph{difficulty:} a number between 0.0 and 5.0 inclusive.  The
difficulty for a particular benchmark will be assigned by running SMT
solvers from the 2010 competition that finished in good
standing and using the formula:

@@

\item \textbf{Retire very easy benchmarks.} %
  The most difficult~300 non-check non-unknown benchmarks in each
  division are always included, together with all benchmarks on which
  at least one 2010 solver required more than 5~seconds.  \emph{This
    is intended to have the effect of retiring ``very easy''
    benchmarks that were solved by every 2010 solver in less
    than 5~seconds, \emph{unless} doing so reduces the pool of
    benchmarks for the division to less than~300.}

\item \textbf{Retire inappropriate benchmarks.} %
  The competition organizers will remove from the eligibility pool
  certain SMT-LIB benchmarks that are inappropriate or uninteresting
  for competition, or cut the size of certain benchmark families to
  avoid their over-representation.

@@

\item\textbf{Category slot allocation.} %
  Next, 200 slots are allocated for the division as follows:\footnote{The
  number ``200'' is a guideline, and is expected to be used.  However, the
  SMT-COMP organizers reserve the right to reduce this to an appropriate
  value to ensure a timely end to the competition, and may do so on a
  per-division basis.  The category allotments, etc., will remain the same
  proportion of the total.}
%
  \begin{itemize}
  \item 170 from category \emph{industrial}
  \item 20 from category \emph{crafted}
  \item 10 from category \emph{random}
  \end{itemize}
%
  If there are fewer than 20 (respectively, 10) \emph{crafted} or
  \emph{random} benchmarks in the division pool, more
  \emph{industrial} slots are allocated to make 200 total for the
  division.  If there are too few \emph{industrial} benchmarks in the
  division pool, more \emph{crafted} slots are allocated to make 200
  total for the division.  (In no division are there not enough of
  both industrial and crafted benchmarks.)

@@

Pseudo-random numbers will be generated using the standard C library
function \texttt{random()}, seeded (using \texttt{srandom()}) with the
sum, modulo $2^{30}$, of the numbers provided in the system
descriptions (see Section~\ref{sec:entrants} above) by all SMT-COMP
entrants other than the organizers.  Additionally, the integer part of
the opening value of the New York Stock Exchange Composite Index on
July~12th
will be added to the other seeding values.  This helps provide transparency,
by guaranteeing that the organizers cannot manipulate the seed in
favor of their submitted solvers.  Benchmarks will also be slightly
scrambled before the competition, using a simple benchmark scrambler %the same scrambler as last year,
seeded with the same seed as the benchmark selector.  Both the
scrambler and benchmark selector will be publicly available before the
competition.  Naturally, solvers must not rely on previously
determined identifying syntactic characteristics of competition
benchmarks in testing satisfiability (violation of this is considered
cheating).

\subsection{Application track}

\header{Benchmark sources.} %
Benchmarks for the application track will be collected by the SMT-COMP organizers.
Any benchmark available to the organizers by May 15th (see the timeline in
Section~\ref{sec:timeline}) will be considered eligible.

\header{Benchmark availability.} 
A first release of the application track
benchmarks will be made available within March 31st, 2011.
More benchmarks can be collected, until May 15th.
No additional
benchmarks will be added after this date, but benchmarks can be
modified or removed to fix possible bugs or other issues. 
The final release that will be used for the competition will be posted on 
% rb 23/3: changed from June 30th to agree with sec 6.
June 15th.

\header{Benchmark demographics and selection.} 
All the available benchmarks will be used for the competition. 
For this first year of the application track, no difficulty will be assigned to the benchmarks.
Benchmarks will be slightly scrambled before the competition, using the same scrambler and 
random seed as the main track. 
% rb 25/3: added
For this year, the selection algorithm for the application track will be chosen after the
deadline of May 15th, when the exact demographic of the benchmarks (collected with a usual
``call for benchmarks'') will be available.

@@

\begin{figure}
\begin{center}
\begin{tabular}{|l|c|c|}
\hline \textbf{Reported} & \textbf{Correct?} &
\textbf{Point/penalty}\\ \hline \texttt{unsat} & yes & +1 \\ \hline
\texttt{unsat} & no & -8 \\ \hline \texttt{sat} & yes & +1 \\ \hline
\texttt{sat} & no & -8 \\ \hline \texttt{unknown} & n.a. & 0 \\ \hline
\emph{timeout} & n.a. & 0 \\ \hline
\emph{abort} & n.a. & 0 \\ \hline
\end{tabular}
\end{center}
\caption{\label{fig:points}Points and Penalties for the main and parallel tracks.}
\end{figure}

Winners in each Problem Division for which there are at least three
entrants from distinct research groups competing will be taken to be
those with the highest score. In addition to recognizing the overall
winner in each division, the top solver that provides its source code will also be
recognized in each division.  
For Problem Divisions with fewer than three
entrants, the results will be reported but no winner officially
declared.

For the main and parallel tracks, the score will be computed according to the system of points and
penalties in Figure~\ref{fig:points}. 
As indicated, correct answers
are awarded a positive number of points, while incorrect answers are
penalized by assigning a negative number of points.  Timeouts, aborts,
and reports of \texttt{unknown} are awarded zero points.
%
In the event of a tie
in total number of points, the solver with the lowest average CPU time
on formulas for which it gave a correct answer will be
considered the winner.

For the application track, the score will be computed according to the procedure described in \S\ref{sec:exec:application}.

@@

\appendix
\section{Sample benchmark scripts for the main track}

>>>>

\title{Satisfiability Modulo Theories Competition (SMT-COMP) 2010: Rules and 
Procedures}

@@

\author{
Clark Barrett\\
Computer Science\\
New York University\\
\\
Albert Oliveras \\
LSI Department \\
Technical University of Catalonia \\
\and
Morgan Deters \\
Computer Science \\
New York University \\
\\
Aaron Stump \\
Department of Computer Science \\
University of Iowa}

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{PSS02,leberre+03}.  More information
on the history and motivation for SMT-COMP can be found at the
SMT-COMP web site, \url{www.smtcomp.org}, and in reports on previous
competitions~(\cite{SMTCOMP-2008,BDOS08,SMTCOMP-FMSD,SMTCOMP-JAR}).
SMT-COMP 2010 is jointly affiliated with the 22nd International
Conference on Computer Aided Verification~(CAV), and with the 13th
International Conference on Theory and Applications of Satisfiability
Testing, both held in Edinburgh, Scotland.

The rest of this document, updated from the last year's version,
describes the rules and competition procedures for SMT-COMP 2009.
The primary changes from last year's version are:

\begin{itemize}
\item adoption of the SMT-LIB 2.0 standard;
\item updates to the benchmark selection algorithm, including a change
  in the notion of what constitutes a ``family'' of benchmarks;
\item updates to the difficulty calculation;
\item a track for concurrent solvers, to run before the
  regular competition.
\item 5 random ``check'' benchmarks will be generated using Robert
  Brummayer's SMT fuzzing tool, after all the solvers are uploaded.
  These will be included in the competition, to help promote attention
  to the problem of solver correctness (cf.~\cite{brummayer+09}).
\end{itemize}

@@

\header{Solver format.} %
An entrant to SMT-COMP is an SMT solver
submitted using the SMT-Exec service.  SMT-Exec enables members of the
SMT research community to run solvers on jobs consisting of benchmarks
from the SMT-LIB benchmark library.  Jobs are run on a 12-node computer
cluster purchased with funds from a National Science Foundation
Computing Research Infrastructure grant.  SMT-Exec is provided free of
charge, but it does require a minimal registration, which verifies an
email address and prevents misuse of the service.  Registered users
may then upload their own solvers to run, or may run public solvers
already uploaded to SMT-Exec.  SMT-Exec provides a variety of tabular
and graphical displays of results of solver executions, for
comparison. For the main (sequential-solver) track of SMT-COMP 2010,
SMT-Exec will be configured so that jobs are run on a 64-bit,
uniprocessor Linux kernel.  For the concurrent-solver track, SMT-Exec
will be configured so that jobs are run on a 64-bit, multiprocessor
Linux kernel.\footnote{This year, the competition may take
  advantage of off-site computing resources not normally part of
  SMT-Exec; please stay tuned to the SMT-COMP mailing list for final
  details regarding the effective SMT-COMP 2010 competition cluster.}

@@

\subsection*{Deadlines} % To guarantee inclusion in the competition,

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via SMT-Exec by 7pm, Eastern U.S.
time, July~8th, 2010.  At that time the SMT-Exec service will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until 7pm, Eastern U.S. time, July~10th, 2010.  We strongly encourage
participants to use this two day grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using SMT-Exec after the original deadline on July~8th.

@@

\paragraph{Concurrent solver demonstration.} %
Solvers employing concurrency are invited to participate in a
demonstration that will be run before the regular competition.
Concurrent solvers must be submitted by the same July~8th deadline,
with the same two-day grace period for resubmission as in the regular
competition.  Submissions to both the concurrent solver track
and the main competition are independent: participants must submit
explicitly to both events to participate in both, and they may submit
different (or differently configured) solvers to each.  Benchmarks may
or may not be scrambled for this division.  Entrants should still
include a system description, as for the regular competition.

\medskip
\noindent
Note that the SMT-COMP organizers are considering ways to include
``historical'' solvers in both the main and concurrent track for
demonstration and comparison.  The organizers reserve the right to
include or exclude such solvers, and to make simple modifications
to historical sequential solvers to (na\"ively) take advantage of
multiple processors.

@@

\header{Dates of competition.} % We anticipate that the bulk of the
competition will take place during the course of CAV 2010, SAT 2010,
and SMT 2010, from July~14th to~19th.  Results will be announced in a
special session of CAV, on the last day of the conference, as well as
on the SMT-COMP web site.  Intermediate results will be regularly
posted to the SMT-COMP website as the competition runs.

\header{Input and Output.} % Each SMT-COMP entrant, when executed,
Participating solvers must read a single benchmark script (defined
below, not part of the 2.0 standard), presented on its standard input
channel.  The script is in the concrete syntax of the SMT-LIB format,
version 2.0.  A benchmark script is essentially just the translation
of a benchmark from the version 1.2 specification.  In more detail, a
\textbf{benchmark script} is just a script where:

\begin{itemize}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command.
\item The $\akey{exit}$ command is the last command.
\item There is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item There is at most one $\akey{set-info}$ command for \texttt{status}.
\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item Extra symbols are declared or defined exactly once before any
  use, using $\akey{declare-sort}$, $\akey{define-sort}$,
  $\akey{declare-fun}$, or $\akey{define-fun}$.  They must be part of
  the allowed signature expansion for the logic.
\item No other commands besides the ones just mentioned may be used.
\end{itemize}

\noindent Note that the final version of the 2.0 format will be frozen
in early March.  The SMT-LIB format specification is publicly
available from the ``Documents'' section of the SMT-LIB
website~\cite{SMT-LIB}.  Solvers will be given formulas just from the
Problem Divisions indicated in their system descriptions.

Each SMT-COMP entrant is then expected to attempt to report on its
standard output channel whether the formula is satisfiable
(``\texttt{sat}'', without the quotation marks) or unsatisfiable
(``\texttt{unsat}'').  An entrant may also report ``\texttt{unknown}''
to indicate that it cannot determine satisfiability of the formula.
For more detailed information on the output format, see the
description on the SMT-Exec ``Upload a Solver'' page.

@@

\header{Aborts and unparsable output.} %
Solvers which exit before the time
limit without reporting a result (\ie\ due to exhausting memory, crashing,
or producing output other than \texttt{sat}, \texttt{unsat}, or
\texttt{unknown})
will be considered to have aborted.  An abort scores the same as if
the output is ``\texttt{unknown}''.

@@

@@

We expect the problem divisions for SMT-COMP 2009 to include the following
SMT-LIB \emph{logics}.  These logics are specified in SMT-LIB format on the
SMT-LIB web page.  Note that the ``QF\_'' prefix means the division's
formulas are quantifier-free.  However, the organizers reserve the right to add
(remove) divisions if (not) enough benchmarks and solvers exist for a
particular division.

@@

\header{Benchmark sources.} %
Benchmark formulas for these divisions
will be drawn from the SMT-LIB library.  Any benchmarks added to
SMT-LIB by the July 1 release (see the timeline in
Section~\ref{sec:timeline}) will be considered eligible.  SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.

\header{Benchmark availability.} % A first release of the competition
Benchmarks will be made available on May 1st, 2010. A second and
almost final release will be available on June 1st. No additional
benchmarks will be added after this date, but benchmarks can be
modified or removed to fix possible bugs or other issues. The final
release that will be used for the competition will be posted on June
30th. The set of selected benchmarks will be published when the
competition begins.

\header{Benchmark demographics.} %
In SMT-LIB, benchmarks are organized according to \emph{families}.  A benchmark
family contains problems that are similar in some significant way.  Typically
they come from the same source or application, or are all output by the same
tool.  \emph{Each top-level subdirectory within a division represents a distinct
family.}  This represents a significant change from previous
competition years, where \emph{terminal} subdirectories delineated families.
%
Each benchmark in SMT-LIB also has a \emph{category}.  There are four possible
categories:
%
\begin{itemize}
\item \emph{check.} These benchmarks are hand-crafted to test whether
  solvers support specific features of each division.  In particular,
  there are checks for integer completeness (\ie\ benchmarks that are
  satisfiable under the reals but not under the integers) and big
  number support (\ie\ benchmarks that are likely to fail if integers
  cannot be represented beyond some maximum value, such as
  $2^{31}-1$).

  \noindent\textbf{New for 2010:} using the same random seed as
  for benchmark selection and scrambling, 5 ``check'' benchmarks will
  be randomly generated using Robert Brummayer's SMT fuzzing
  tool~\cite{brummayer+09}.  The (exact version of this) tool will be
  publicly available from the SMT 2010 web site before the
  competition.  Since these benchmarks are generated after the random
  seed is fixed, they cannot be known in advance to any competitors,
  including the organizers.  The rationale for including these
  benchmarks is to try to take a step towards stronger certification
  that solvers are correct.  In future years, we envision SMT-COMP
  requiring that solvers pass some kind of pre-qualifying round based
  on SMT fuzzing.  The inclusion of these fuzzing benchmarks that are
  not known in advance is intended to encourage (without requring)
  solver implementors to test their solvers using fuzzing or similar
  bug-discovery techniques.
\item \emph{industrial.} These benchmarks come from some real application and are produced by tools such as bounded model checkers, static analyzers, extended static checkers, etc.
\item \emph{random.} These benchmarks are randomly generated.
\item \emph{crafted.} This category is for all other benchmarks.  Usually,
  benchmarks in this category are designed to be particularly difficult or to
  test a specific feature of the logic.
\end{itemize}

\header{Benchmark selection.} %
Before the selection process, each benchmark will be assigned a
\emph{difficulty:} a number between 0.0 and 5.0 inclusive.  The
difficulty for a particular benchmark will be assigned by running SMT
solvers from the 2009 competition that finished in good
standing\footnote{Note that the corrected version of MathSat that was
  submitted after the 2009 deadline, which ran \textit{hors concours}
  in the 2009 competition, will be used instead of its ``official''
  counterpart.} and using the formula:

@@

\item \textbf{Retire very easy benchmarks.} %
  The most difficult~300 non-check non-unknown benchmarks in each
  division are always included, together with all benchmarks on which
  at least one 2009 solver required more than 5~seconds.  \emph{This
    is intended to have the effect of retiring ``very easy''
    benchmarks that were solved by every 2009 solver in less
    than 5~seconds, \emph{unless} doing so reduces the pool of
    benchmarks for the division to less than~300.}

@@

\item\textbf{Category slot allocation.} %
  Next, 200 slots are allocated for the division as follows:\footnote{The
  number "200" is a guideline, and is expected to be used.  However, the
  SMT-COMP organizers reserve the right to reduce this to an appropriate
  value to ensure a timely end to the competition, and may do so on a
  per-division basis.  The category allotments, etc., will remain the same
  proportion of the total.}
%
  \begin{itemize}
  \item 170 from category \emph{industrial}
  \item 20 from category \emph{crafted}
  \item 10 from category \emph{random}
  \end{itemize}
%
  If there are fewer than 20 (respectively, 10) \emph{crafted} or
  \emph{random} benchmarks in the division pool, more
  \emph{industrial} slots are allocated to make 200 total for the
  division.  If there are too few \emph{industrial} benchmarks in the
  division pool, more \emph{crafted} slots are allocated to make 200
  total for the division.  (In no division are there not enough of
  both industrial and crafted benchmarks.)

@@

Pseudo-random numbers will be generated using the standard C library
function \texttt{random()}, seeded (using \texttt{srandom()}) with the
sum, modulo $2^{30}$, of the numbers provided in the system
descriptions (see Section~\ref{sec:entrants} above) by all SMT-COMP
entrants other than the organizers.  Additionally, the integer part of
the opening value of the New York Stock Exchange Composite Index on August 4th
will be added to the other seeding values.  This helps provide transparency,
by guaranteeing that the organizers cannot manipulate the seed in
favor of their submitted solvers.  Benchmarks will also be slightly
scrambled before the competition, using the same scrambler as last year,
seeded with the same seed as the benchmark selector.  Both the
scrambler and benchmark selector will be publicly available before the
competition.  Naturally, solvers must not rely on previously
determined identifying syntactic characteristics of competition
benchmarks in testing satisfiability (violation of this is considered
cheating).

@@

\begin{figure}
\begin{center}
\begin{tabular}{|l|c|c|}
\hline \textbf{Reported} & \textbf{Correct?} &
\textbf{Point/penalty}\\ \hline \texttt{unsat} & yes & +1 \\ \hline
\texttt{unsat} & no & -8 \\ \hline \texttt{sat} & yes & +1 \\ \hline
\texttt{sat} & no & -8 \\ \hline \texttt{unknown} & n.a. & 0 \\ \hline
\emph{timeout} & n.a. & 0 \\ \hline
\emph{abort} & n.a. & 0 \\ \hline
\end{tabular}
\end{center}
\caption{\label{fig:points}Points and Penalties}
\end{figure}

Winners in each Problem Division for which there are at least three
entrants from distinct research groups competing will be taken to be
those with the highest score, according to the system of points and
penalties in Figure~\ref{fig:points}.  In addition to recognizing the overall
winner in each division, the top solver providing its source code will also be
recognized in each division.  As indicated, correct answers
are awarded a positive number of points, while incorrect answers are
penalized by assigning a negative number of points.  Timeouts, aborts,
and reports of \texttt{unknown} are awarded zero points.
%
In the event of a tie
in total number of points, the solver with the lowest average CPU time
on formulas for which it gave a correct answer will be
considered the winner.  For Problem Divisions with fewer than three
entrants, the results will be reported but no winner officially
declared.

@@
